---
title: "Spatial Analysis- methodology revisited 2"
author: "Paola Kim-Blanco"
date: "April 23, 2016"
output: html_document
---

```{r setup, include=FALSE}
library(fmsb)
library(Rcmdr)
library(maptools)
library(rgdal)
library(spdep)
library(ctv)
library(knitr)
library(rmarkdown)
library(lmtest)
library(openxlsx)
library(MASS)
library(foreign)
library(fBasics)
library(robust)
library(spgwr)
library(aod)

```

#Methodology

1- Data transformations, preliminary
2- VIF test for all variables
3- Descriptives of relevant variables (from previous runs)
4- OLS + conduct Breusch -Pagan test for heteroskedasticity
5- Look at residuals, run outlier analysis.
6- OLS without outliers + BP test again
7- OLS robust + BP test again
8- Moran's I on weighted residuals
9- Lagrange tests to select spatial model
10- Spatial model
11- Moran's I on residuals for sp model--is it the correct model?


###1- Data transformations:
```{r }

x <- readOGR(dsn = "/Volumes/G-DRIVE mobile/Thesis/_2Data/gis/master_db", layer = "master_db7", verbose = FALSE)
x@data$ALLHAZARDS <- x@data$EQK + x@data$FLOOD + x@data$FROST + x@data$HEATWAVE + x@data$RAIN + x@data$HURR

#log transformations
#even though ln(0) = Inf, I am transforming to zero because 0 income or 0 events is 0.
attach(x@data)
x@data$ln_INRATE <- ifelse(INRATE == 0, 0, log(INRATE))
x@data$ln_OUTRATE <- ifelse(OUTRATE == 0, 0, log(OUTRATE))
x@data$ln_INCOME <- ifelse(INCOME == 0, 0, log(INCOME))
x@data$ln_VDEATH <- ifelse(VDEATH == 0, 0, log(VDEATH))
x@data$ln_FLOOD <- ifelse(FLOOD == 0, 0, log(FLOOD))
x@data$ln_FROST <- ifelse(FROST == 0, 0, log(FROST))
x@data$ln_HURR <- ifelse(HURR == 0, 0, log(HURR))
x@data$ln_RAIN <- ifelse(RAIN == 0, 0, log(RAIN))
x@data$ln_EQK <- ifelse(EQK == 0, 0, log(EQK))
x@data$ln_ALLHAZARDS <- ifelse(ALLHAZARDS == 0, 0, log(ALLHAZARDS))
x@data$ln_ECONEST <- ifelse(ECONEST == 0, 0, log(ECONEST))
x@data$ln_EXPSCHOOL <- ifelse(EXPSCHOOLI == 0, 0, log(EXPSCHOOLI))
x@data$ln_PINV <- ifelse(PINV == 0, 0, log(PINV))
x@data$ln_PURBAN <- ifelse(PURBAN == 0, 0, log(PURBAN))
detach(x@data)
```


###2- Run VIF
```{r}
#run vif function-- got it from https://beckmw.wordpress.com/2013/02/05/collinearity-and-stepwise-vif-selection/

vif_func<-function(in_frame,thresh=10,trace=T,...){
    
    require(fmsb)
    
    if(class(in_frame) != 'data.frame') in_frame<-data.frame(in_frame)
    
    #get initial vif value for all comparisons of variables
    vif_init<-NULL
    var_names <- names(in_frame)
    for(val in var_names){
        regressors <- var_names[-which(var_names == val)]
        form <- paste(regressors, collapse = '+')
        form_in <- formula(paste(val, '~', form))
        vif_init<-rbind(vif_init, c(val, VIF(lm(form_in, data = in_frame, ...))))
    }
    vif_max<-max(as.numeric(vif_init[,2]))
    
    if(vif_max < thresh){
        if(trace==T){ #print output of each iteration
            prmatrix(vif_init,collab=c('var','vif'),rowlab=rep('',nrow(vif_init)),quote=F)
            cat('\n')
            cat(paste('All variables have VIF < ', thresh,', max VIF ',round(vif_max,2), sep=''),'\n\n')
        }
        return(var_names)
    }
    else{
        
        in_dat<-in_frame
        
        #backwards selection of explanatory variables, stops when all VIF values are below 'thresh'
        while(vif_max >= thresh){
            
            vif_vals<-NULL
            var_names <- names(in_dat)
            
            for(val in var_names){
                regressors <- var_names[-which(var_names == val)]
                form <- paste(regressors, collapse = '+')
                form_in <- formula(paste(val, '~', form))
                vif_add<-VIF(lm(form_in, data = in_dat, ...))
                vif_vals<-rbind(vif_vals,c(val,vif_add))
            }
            max_row<-which(vif_vals[,2] == max(as.numeric(vif_vals[,2])))[1]
            
            vif_max<-as.numeric(vif_vals[max_row,2])
            
            if(vif_max<thresh) break
            
            if(trace==T){ #print output of each iteration
                prmatrix(vif_vals,collab=c('var','vif'),rowlab=rep('',nrow(vif_vals)),quote=F)
                cat('\n')
                cat('removed: ',vif_vals[max_row,1],vif_max,'\n\n')
                flush.console()
            }
            
            in_dat<-in_dat[,!names(in_dat) %in% vif_vals[max_row,1]]
            
        }
        
        return(names(in_dat))
        
    }
    
}

#all data
# took out heatwave because it was causing problems
y1 <- x@data[, c(32, 34:47, 18:22)]
vif_func(in_frame = y1, thresh = 5, trace = FALSE)

#selected data: natural disasters separately
y2 <- x@data[, c(32, 34:47)]
vif_func(in_frame = y2, thresh = 5, trace = F)

#selected data: total natural disasters
y3 <- x@data[, c(32, 34:37, 43:47)]
vif_func(in_frame = y3, thresh = 5, trace = F)

#selected data: % localities with reported damages, by type of disaster
y4 <- x@data[, c(32, 34:37, 44:47,18:22)]
vif_func(in_frame = y4, thresh = 5, trace = F)


```


###3- Descriptives of relevant variables (from previous runs)
```{r}
attach(x@data)
hist(ln_INRATE)
hist(ln_OUTRATE)
hist(ln_INCOME)
hist(INCOME)
hist(ln_EXPSCHOOL)
hist(EXPSCHOOLI)
hist(VDEATH)
hist(ln_VDEATH)
hist(ln_PURBAN)
hist(ln_PINV)
hist(PINV)
hist(ECONEST)
hist(ln_ECONEST)

hist(ln_FLOOD)
hist(ln_FROST)
hist(ln_EQK)
hist(ln_HURR)
hist(ln_RAIN)
hist(ln_ALLHAZARDS)

hist(FLOOD)
hist(FROST)
hist(EQK)
hist(HURR)
hist(RAIN)
hist(ALLHAZARDS)

hist(PFROST)
hist(PDROUGHT)
hist(PEQK)
hist(PHURR)
hist(PFLOOD)

```




###4- OLS + conduct Breusch -Pagan test for heteroskedasticity

```{r}
#ln_INRATE
#ln_OUTRATE
#INCOME
#EXPSCHOOLI
#ln_VDEATH
#ln_PURBAN
#ln_PINV
#ln_ECONEST

#In-rate

###All hazards, por separado
summary(in1 <- stepwise(lm(ln_INRATE ~ RURDUMMY + INCOME + ln_VDEATH + ln_ECONEST + ln_PINV + EXPSCHOOLI + ln_PURBAN + ln_FLOOD + ln_FROST + ln_HURR + ln_RAIN + ln_EQK , data =  x)))
bptest(in1) #variables checked against multicolinearity/ yes, heteroskedasticity, indiv hazards not significant

summary(in2 <- lm(ln_INRATE ~ INCOME + ln_ECONEST + ln_PINV + EXPSCHOOLI + ln_FLOOD + ln_FROST + ln_HURR + ln_RAIN + ln_EQK , data =  x))
bptest(in2) #yes, heteroskedasticity, only rain significant


###All hazards, sum
summary(in3 <- stepwise(lm(ln_INRATE ~ RURDUMMY + INCOME + ln_VDEATH + ln_ECONEST + ln_PINV + EXPSCHOOLI + ln_ALLHAZARDS , data =  x)))
bptest(in3) #variables checked against multicolinearity/ yes, heteroskedasticity, all hazards not significant

summary(in4 <- lm(ln_INRATE ~ INCOME + ln_ECONEST + ln_PINV + EXPSCHOOLI + ln_ALLHAZARDS , data =  x))
bptest(in4) #yes, heteroskedasticity, all hazards not significant


#Percent rural villages with damages
summary(in5 <- stepwise(lm(ln_INRATE ~ RURDUMMY + INCOME + ln_VDEATH + ln_ECONEST + ln_PINV + EXPSCHOOLI + PDROUGHT + PFROST + PFLOOD + PEQK + PHURR , data =  x)))
bptest(in5) #variables checked against multicolinearity/ yes, heteroskedasticity, all % village damages not significant

summary(in6 <- lm(ln_INRATE ~ INCOME + ln_ECONEST + ln_PINV + EXPSCHOOLI + PDROUGHT + PFROST + PFLOOD + PEQK + PHURR , data =  x))
bptest(in6) #yes, heteroskedasticity, all % village damages not significant

```

```{r}

#Out-rate
###All hazards, por separado
summary(out1 <- stepwise(lm(ln_OUTRATE ~ RURDUMMY + INCOME + ln_VDEATH + ln_ECONEST + ln_PINV + EXPSCHOOLI + ln_PURBAN + ln_FLOOD + ln_FROST + ln_HURR + ln_RAIN + ln_EQK , data =  x)))
bptest(out1) #variables checked agains multicolinearity; heteroskedastic; only frost is significant

summary(out2 <- lm(ln_OUTRATE ~ INCOME + ln_ECONEST + ln_PINV + EXPSCHOOLI + ln_FLOOD + ln_FROST + ln_HURR + ln_RAIN + ln_EQK , data =  x))
bptest(out2) #heteroskedastic; rain,hurricanes and frost significant

#All hazards, sum
summary(out3 <- stepwise(lm(ln_OUTRATE ~ RURDUMMY + INCOME + ln_VDEATH + ln_ECONEST + ln_PINV + EXPSCHOOLI + ln_ALLHAZARDS , data =  x)))
bptest(out3) #variables checked against multicolinearity/ yes, heteroskedasticity, all hazards significant

summary(out4 <- lm(ln_OUTRATE ~ INCOME + ln_ECONEST + ln_PINV + EXPSCHOOLI + ln_ALLHAZARDS , data =  x))
bptest(out4) #yes, heteroskedasticity, all hazards significant


#Percent rural villages with damages
summary(out5 <- stepwise(lm(ln_OUTRATE ~ RURDUMMY + INCOME + ln_VDEATH + ln_ECONEST + ln_PINV + EXPSCHOOLI + PDROUGHT + PFROST + PFLOOD + PEQK + PHURR , data =  x)))
bptest(out5) #variables checked against multicolinearity/ yes, heteroskedasticity, %drought, %frost, %flood, %hurr are significant

summary(out6 <- lm(ln_OUTRATE ~ INCOME + ln_ECONEST + ln_PINV + EXPSCHOOLI + PDROUGHT + PFROST + PFLOOD + PEQK + PHURR , data =  x))
bptest(out6) #yes, heteroskedasticity/ everything is significant except % earthquakes
```

###5- Look at residuals, adjust models

```{r}
residualPlots(in2) #residual plots
avPlots(in2, id.n=2, id.cex=0.7) #influential observations
qqPlot(in2, id.n=3)
outlierTest(in2)
influenceIndexPlot(in2, id.n = 3) 

residualPlots(in4) #residual plots
avPlots(in4, id.n=2, id.cex=0.7) #influential observations
qqPlot(in4, id.n=3)
outlierTest(in4)
influenceIndexPlot(in4, id.n = 3) 
#from looking at added-variable plots, took out ln_PINV, and run again

summary(in4 <- lm(ln_INRATE ~ INCOME + ln_ECONEST + EXPSCHOOLI + ln_ALLHAZARDS , data =  x))
bptest(in4) #yes, heteroskedasticity, all hazards not significant

residualPlots(in4) #residual plots
avPlots(in4, id.n=2, id.cex=0.7) #influential observations
qqPlot(in4, id.n=3)
outlierTest(in4)
influenceIndexPlot(in4, id.n = 3) 

residualPlots(in6) #residual plots
avPlots(in6, id.n=2, id.cex=0.7) #influential observations
qqPlot(in6, id.n=3)
outlierTest(in6)
influenceIndexPlot(in6, id.n = 3) 

#from residual plot, take out ln_PINV and use ln_INCOME
summary(in6 <- lm(ln_INRATE ~ ln_INCOME + ln_ECONEST + EXPSCHOOLI + PDROUGHT + PFROST + PFLOOD + PEQK + PHURR , data =  x))
bptest(in6)

residualPlots(in6) #residual plots
avPlots(in6, id.n=2, id.cex=0.7) #influential observations
qqPlot(in6, id.n=3)
outlierTest(in6)
influenceIndexPlot(in6, id.n = 3)


###out

residualPlots(out2)
avPlots(out2, id.n=2, id.cex=0.7) #influential observations
qqPlot(out2, id.n=3)
outlierTest(out2)
influenceIndexPlot(out2, id.n = 3)

#use ln_INCOME, ln_PINV, take out ln_HURR, ln_FLOOD, ln_EQK
summary(out2 <- lm(ln_OUTRATE ~ ln_INCOME + ln_ECONEST + EXPSCHOOLI + ln_FROST + ln_RAIN , data =  x))
bptest(out2)
# lots of outliers, same one (918) as in

residualPlots(out4)
avPlots(out4, id.n=2, id.cex=0.7) #influential observations
qqPlot(out4, id.n=3)
outlierTest(out4)
influenceIndexPlot(out4, id.n = 3)
#take out ln_PINV, use ln_INCOME

summary(out4 <- lm(ln_OUTRATE ~ ln_INCOME + ln_ECONEST + EXPSCHOOLI + ln_ALLHAZARDS , data =  x))
bptest(out4)
#lost of outliers

## correcting for income and ln_PINV
summary(out6 <- lm(ln_OUTRATE ~ ln_INCOME + ln_ECONEST  + EXPSCHOOLI + PDROUGHT + PFROST + PFLOOD + PEQK + PHURR , data =  x))
bptest(out6)

residualPlots(out6)
avPlots(out6, id.n=2, id.cex=0.7) #influential observations
qqPlot(out6, id.n=3)
outlierTest(out6)
influenceIndexPlot(out6, id.n = 3)

```

###6- Remove outliers

```{r}

outlierTest(in2)
outlierTest(in4)
outlierTest(in6)
outlierTest(out2)
outlierTest(out4)
outlierTest(out6)

outliers <- data.frame(mun_id = c("07113", "20319", "07056", "19041", "23010", "17007", "07022", "07118","20053", "20233", "15012", "20552"))

y <- subset(x, !mun_id %in% outliers$mun_id)
```

###7- OLS with no outliers
```{r}
summary(in2 <- lm(ln_INRATE ~ ln_INCOME + ln_ECONEST + EXPSCHOOLI + ln_FLOOD + ln_FROST + ln_HURR + ln_RAIN + ln_EQK , data =  y))
bptest(in2)

summary(in4 <- lm(ln_INRATE ~ ln_INCOME + ln_ECONEST + EXPSCHOOLI + ln_ALLHAZARDS , data =  y))
bptest(in4)




summary(out2 <- lm(ln_OUTRATE ~ ln_INCOME + ln_ECONEST + EXPSCHOOLI + ln_FROST + ln_RAIN , data =  y))
bptest(out2)

summary(out4 <- lm(ln_OUTRATE ~ ln_INCOME + ln_ECONEST + EXPSCHOOLI + ln_ALLHAZARDS , data =  y))
bptest(out4)


#all models are still heteroskedastic

##models #in6 and out6 need to have a separate run because they only pertain rural areas---
w <- y[y@data$RURDUMMY == 1,]

summary(in6 <- lm(ln_INRATE ~ ln_INCOME + ln_ECONEST + EXPSCHOOLI + PDROUGHT + PFROST + PFLOOD + PEQK + PHURR , data =  w))
bptest(in6)
logLik(in6)

summary(out6 <- lm(ln_OUTRATE ~ ln_INCOME + ln_ECONEST  + EXPSCHOOLI + PDROUGHT + PFROST + PFLOOD + PEQK + PHURR , data =  w))
bptest(out6)
logLik(out6)


```

###8- Run ronust OLS

```{r}
summary(in2r <- rlm(ln_INRATE ~ ln_INCOME + ln_ECONEST + EXPSCHOOLI + ln_FLOOD + ln_FROST + ln_HURR + ln_RAIN + ln_EQK , data =  y))
bptest(in2r)

summary(in4r <- rlm(ln_INRATE ~ ln_INCOME + ln_ECONEST + EXPSCHOOLI + ln_ALLHAZARDS , data =  y))
bptest(in4r)

summary(in6r <- rlm(ln_INRATE ~ ln_INCOME + ln_ECONEST + EXPSCHOOLI + PDROUGHT + PFROST + PFLOOD + PEQK + PHURR , data =  y))
bptest(in6r)

summary(out2r <- rlm(ln_OUTRATE ~ ln_INCOME + ln_ECONEST + EXPSCHOOLI + ln_FROST + ln_RAIN , data =  y))
bptest(out2r)

summary(out4r <- rlm(ln_OUTRATE ~ ln_INCOME + ln_ECONEST + EXPSCHOOLI + ln_ALLHAZARDS , data =  y))
bptest(out4r)

summary(out6r <- rlm(ln_OUTRATE ~ ln_INCOME + ln_ECONEST  + EXPSCHOOLI + PDROUGHT + PFROST + PFLOOD + PEQK + PHURR , data =  y))
bptest(out6r)

```

BP tests are still significant in all cases, therefore models are still heteroskedastic. This is indicative of
patterns in the error which can be modelled with spatial models. Will use OLS for spatial modeling

###9- Moran's I on robust or linear?

```{r}
queen1 <- poly2nb(y)
queen_w <- nb2listw(queen1, style = "W", zero.policy = TRUE)
queen1rural <- poly2nb(w)
queen_wrural <- nb2listw(queen1rural, style = "W", zero.policy = TRUE)

y@data$in2_res <- in2r$residuals
moran.test(y@data$in2_res, queen_w, randomisation = TRUE, alternative = "two.sided")
y@data$in4_res <- in4$residuals
moran.test(y@data$in4_res, queen_w, randomisation = TRUE, alternative = "two.sided")
w@data$in6_res <- in6$residuals
moran.test(w@data$in6_res, queen_wrural, randomisation = TRUE, alternative = "two.sided")

y@data$out2_res <- out2$residuals
moran.test(y@data$out2_res, queen_w, randomisation = TRUE, alternative = "two.sided")
y@data$out4_res <- out4$residuals
moran.test(y@data$out4_res, queen_w, randomisation = TRUE, alternative = "two.sided")
w@data$out6_res <- out6$residuals
moran.test(w@data$out6_res, queen_wrural, randomisation = TRUE, alternative = "two.sided")

AIC(in2, in4, in6, out2, out4, out6)
rmse <- function(error)
{
    sqrt(mean(error^2))
}
rmse(residuals.lm(in2))
rmse(residuals.lm(in4))
rmse(residuals.lm(in6))
rmse(residuals.lm(out2))
rmse(residuals.lm(out4))
rmse(residuals.lm(out6))


```
Very similar results with rmse
Models in4 and out4 have the smallest AIC

###10- Lagrange tests to select spatial model
```{r}
lm.LMtests(in2, queen_w, test = "all") #spatial error
lm.LMtests(in4, queen_w, test = "all") #spatial error
#lm.LMtests(in6, queen_w, test = "all") #spatial error
lm.LMtests(out2, queen_w, test = "all") #spatial error
lm.LMtests(out4, queen_w, test = "all") #spatial lag
#lm.LMtests(out6, queen_w, test = "all") #spatial lag

```


###10- Spatial model

```{r}
summary(in2sp <- errorsarlm(in2, data = y, queen_w))
bptest.sarlm(in2sp)

summary(in4sp <- errorsarlm(in4, data = y, queen_w ))
bptest.sarlm(in4sp)

#summary(in6sp <- errorsarlm(in6, data = y, queen_w))
#bptest.sarlm(in6sp)

summary(out2sp <- errorsarlm(out2, data = y, queen_w))
bptest.sarlm(out2sp)

summary(out4sp <- lagsarlm(out4, data = y, queen_w))
bptest.sarlm(out4sp)

#summary(out6sp <- lagsarlm(out6, data = y, queen_w))
#bptest.sarlm(out6sp)

#writeOGR(y, dsn = "/Volumes/G-DRIVE mobile/Thesis/_2Data/gis/master_db", #layer = "master_db8", driver = "ESRI Shapefile")


```


###11- Moran's I on residuals for sp model

```{r}

moran.test(in2sp$residuals, queen_w, randomisation = TRUE, alternative = "two.sided")
moran.test(in4sp$residuals, queen_w, randomisation = TRUE, alternative = "two.sided")
moran.test(in6sp$residuals, queen_w, randomisation = TRUE, alternative = "two.sided")
moran.test(out2sp$residuals, queen_w, randomisation = TRUE, alternative = "two.sided")
moran.test(out4sp$residuals, queen_w, randomisation = TRUE, alternative = "two.sided")
moran.test(out6sp$residuals, queen_w, randomisation = TRUE, alternative = "two.sided")


```

###12- GWR

```{r}
bw_in2 <- gwr.sel(ln_INRATE ~ ln_INCOME + EXPSCHOOLI + ln_ECONEST + ln_ALLHAZARDS, data = y, adapt = TRUE) 
gwr_in2 <- gwr(ln_INRATE ~ ln_INCOME + EXPSCHOOLI + ln_ECONEST + ln_ALLHAZARDS, data = y, adapt = bw_in2, hatmatrix = TRUE, predict = TRUE, se.fit = TRUE)
gwr_in2$results$rss
gwr_in2$results$AICh
moran.test(gwr_in2$SDF$gwr.e, queen_w)


bw_in4 <- gwr.sel(in4, data = y, adapt = TRUE)
gwr_in4 <- gwr(in4, data = y, adapt = bw_in4, hatmatrix = TRUE, predict = TRUE, se.fit = TRUE)
gwr_in4$results$rss
gwr_in4$results$AICh
moran.test(gwr_in4$SDF$gwr.e, queen_w)


bw_out2 <- gwr.sel(out2, data = y, adapt = TRUE) 
gwr_out2 <- gwr(out2, data = y, adapt = bw_out2, hatmatrix = TRUE, predict = TRUE, se.fit = TRUE)
gwr_out2$results$rss
gwr_out2$results$AICh
moran.test(gwr_out2$SDF$gwr.e, queen_w)

bw_out4 <- gwr.sel(out4, data = y, adapt = TRUE) 
gwr_out4 <- gwr(out4, data = y, adapt = bw_out4, hatmatrix = TRUE, predict = TRUE, se.fit = TRUE)
gwr_out4$results$rss
gwr_out4$results$AICh
moran.test(gwr_out4$SDF$gwr.e, queen_w)


```

### Residuals
```{r}

rmse(residuals.sarlm(in4sp))
rmse(residuals.sarlm(out4sp))
```


### Ploting queen1
```{r}

coords <- coordinates(y)
plot(y)
plot(queen_w, coords, add = TRUE)




```

